### 2023-02-22 paper-thesis/chapters/01-introduction.tex
Improve the introduction a bit

Type checking is commonly done by writing a type checker in a general purpose language. This requires a lot of effort, and it is easy to make mistakes. In this thesis we will specifically focus on dependently typed languages, which differ from other languages because they allow types to be parameterized by values~\cite{lambda_cube}. This allows types to express properties of values that cannot be expressed in a simple type system, such as the length of a list or the well-formedness of a binary search tree. This expressiveness also makes dependent type systems more complicated to type check, since deciding equality of types requires normalization of the terms they are parameterized by~\cite{Coquand_Huet_1988}.

Instead of writing the type checker in a general purpose language, in this thesis we will systematically derive a type checker from a high-level, declarative specification. This allows for a cleaner implementation that is easy to extend and maintain.

We will do this using the Spoofax language workbench, which is a collection of tools that can derive a parser and type checker from a high level specification of the language~\cite{spoofax}. When working with the Spoofax workbench, the Statix meta-language can be used for the specification of static semantics~\cite{scopes_as_types}. It is a declarative language that uses inference rules to define static semantics. Statix aims to cover a broad range of languages and type systems. However, no attempts have been made yet to express a dependently typed language in Statix until now.

\section*{Contributions}
This goal of this thesis is to investigate how well Statix is fit for the task of defining a dependently-typed language. We want to investigate whether typical features of dependently typed languages can be encoded in Statix. The goal is not only to show that Statix can implement it, but also that the implementation is concise.

The technical contributions of this thesis are the following:
\begin{itemize}
	\item We present a implementation of the Calculus of Constructions, a lambda calculus with dependent types, in Statix (Chapter~\ref{chap:baselang})
	\item We show the language is easily extendable, by extending it with booleans among other features (Chapter~\ref{chap:bools})
	\item We show how to add inference of implicit arguments to the implementation (Chapter~\ref{ch:inference})
	\item We show how to add support for inductive datatypes to the implementation (Chapter~\ref{ch:datatypes})
	\item We show how to add support for universes to the implementation (Chapter~\ref{ch:universes})
\end{itemize}

>>>>

While we keep building more and more complex programming languages, their type checkers are still often written in general purpose languages. This takes a lot of effort, and it is easier to make mistakes. Instead of writing the type checker in a general purpose language, in this thesis we will systematically derive a type checker from a high-level, declarative specification. This allows for a cleaner implementation that is easy to extend and maintain.

In this thesis we will specifically focus on dependently typed languages, which differ from other languages because they allow types to be parameterized by values~\cite{lambda_cube}. This allows types to express properties of values that cannot be expressed in a simple type system, such as the length of a list or the well-formedness of a binary search tree. This expressiveness also makes dependent type systems more complicated to type check, since deciding equality of types requires evaluation of the terms they are parameterized by~\cite{Coquand_Huet_1988}.

We will write the declarative specification in the Spoofax language workbench, which is a collection of tools that can derive a parser and type checker from a high level specification of the language~\cite{spoofax}. When working with the Spoofax workbench, the Statix meta-language can be used for the specification of static semantics~\cite{scopes_as_types}. It is a declarative language that uses inference rules and scope graphs~\cite{scope_graphs} to define static semantics. Statix aims to cover a broad range of languages and type systems. However, no attempts have been made yet to express a dependently typed language in Statix until now.

\section*{Contributions}
This goal of this thesis is to investigate how well Statix is fit for the task of defining a dependently-typed language. The goal is not only to show that Statix can implement it, but also that the implementation is more concise than in a general purpose language. We will start by implementing the Calculus of Constructions~\cite{Coquand_Huet_1988}, a lambda calculus with dependent types, as this is one of the most simple dependently typed languages (chapter~\ref{chap:baselang}).

However, this language is not very practical to write actual complex programs in, so to improve this situation we will extend the language with typical features of dependently typed languages. We will do this in the following steps:
\begin{itemize}
	\item We show the language is easily extendable, by extending it with booleans among other features (Chapter~\ref{chap:bools})
	\item We show how to add inference of implicit arguments to the implementation (Chapter~\ref{ch:inference})
	\item We show how to add support for inductive datatypes to the implementation (Chapter~\ref{ch:datatypes})
	\item We show how to add support for universes to the implementation (Chapter~\ref{ch:universes})
\end{itemize}
